### 2018-10-29 variability/var.tex
start

% var.tex : a paper about EPRV in the presence of spectral variability
% This file is part of the EPRV project.
% Copyright 2018 the authors.

% Style notes
% -----------
% - Re-define acronyms in each section? So less flippy flippy?

% To-do items
% -----------
% - figure out what experiments to do and do them?
% - Draft it.
% - Get a 6th keyword; re-consider keywords.

\documentclass[modern]{aastex62}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[sort&compress]{natbib}
\usepackage[hang,flushmargin]{footmisc}

% units macros
\newcommand{\unit}[1]{\mathrm{#1}}
\newcommand{\km}{\unit{km}}
\newcommand{\m}{\unit{m}}
\newcommand{\s}{\unit{s}}
\newcommand{\kms}{\km\,\s^{-1}}
\newcommand{\ms}{\m\,\s^{-1}}
\newcommand{\ang}{\text{\normalfont\AA}}

% math macros
\newcommand{\dd}{\mathrm{d}}
\newcommand{\T}{^{\mathsf{T}}}

% text macros
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\sectionname}{Section}
\newcommand{\todo}[1]{\textcolor{red}{#1}}  % gotta have \usepackage{xcolor} in main doc or this won't work
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\foreign}[1]{\textsl{#1}}
\newcommand{\HARPS}{\project{\acronym{HARPS}}}
\newcommand{\HIRES}{\project{\acronym{HIRES}}}
\newcommand{\RV}{\acronym{RV}}
\newcommand{\CRLB}{\acronym{CRLB}}

\setlength{\parindent}{1.4em} % trust in Hogg
\shorttitle{precise radial velocities for variable stars}
\shortauthors{hogg and others}

\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust in Hogg
\graphicspath{ {figures/} }
\DeclareGraphicsExtensions{.pdf,.eps,.png}

\title{\textbf{Precise radial-velocity measurements in the presence of stellar variability}}

\author[0000-0003-2866-9403]{David W. Hogg}
\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
\affil{Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011, USA}
\affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg}

\begin{abstract}\noindent
% context
foo.
% aims
bar.
% methods
blah.
% results
duh.
\end{abstract}

\keywords{
% atmospheric effects
% ---
binaries: spectroscopic
---
methods: data analysis
---
methods: statistical
---
techniques: radial velocities
---
planets and satellites: detection
}

\section{Introduction}

Hello World!

\acknowledgements
It is a pleasure to thank
  Ben Pope (Flatiron)
and the participants in the weekly Stars Group Meeting at the Flatiron
Institute for valuable contributions to this project.

\todo{...add Hogg grant numbers}

\todo{...add facilities tags}

\todo{...add software tags}

\bibliographystyle{apj}
\bibliography{}%general,myref,inprep}

\end{document}